Data Exploration: IUCN and AquaMaps

### 2016-08-25 clip_fao/fao_clip_pac.Rmd
updated FAO clipping code for mapping and whatnot

library(ggplot2)
library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

@@

am_spp_latlong <- am_spp_latlong %>%
  filter(clongadj >= 30 & clongadj <= 290)

@@

``` {r northmost_band}
### 5N north     - look for spike at 175 W (185E)
am_spp_north <- am_spp_latlong %>%
  filter(centerlat >= 5) %>%
  filter(clongadj <= 260) ### west of North America

am_spp_north_lims <- am_spp_north %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(30, 260), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

@@

am_spp_central_lims <- am_spp_central %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(30, 260), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

@@

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(77, 185, 240), color = 'red')

``` {r southmost_band}
### 60 S to 25 S - look for spikes at 120W (240E), 150E, 80E
am_spp_south <- am_spp_latlong %>%
  filter(centerlat >= -60 & centerlat <= -25)

am_spp_south_lims <- am_spp_south %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(30, 260), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_south_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(80, 150, 240), color = 'red')

``` {r all_bands}

am_spp_all_lims <- am_spp_latlong %>%
  filter(clongadj <= 260) %>% ### west of North America
  filter(clongadj >= 50) %>% ### east of West Africa
  filter(centerlat >= -25) %>% ### north of Australia
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(50, 260), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_all_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(80, 150, 240), color = 'red') +
  geom_vline(xintercept = c(77, 185), color = 'red')

``` {r all_bands_low_data}

datapoor <- spp_quad_list %>%
  filter(is.na(reviewed)) %>%
  filter(occurcells < 50)

am_spp_dp_lims <- am_spp_latlong %>%
  filter(am_sid %in% datapoor$am_sid) %>%
  filter(clongadj <= 260) %>% ### west of North America
  filter(clongadj >= 50) %>% ### east of West Africa
  filter(centerlat >= -25) %>% ### north of Australia
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(50, 260), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_dp_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(80, 240), color = 'red') +
  geom_vline(xintercept = c(77, 185), color = 'red')

@@

### this first batch is Coral Triangle-area species
clipped_spp1 <- am_spp_all_lims %>%
  # filter(long %in% c(77, 185)) %>%
  filter((long == 77  & lim == 'w_lim') | 
         (long == 185 & lim == 'e_lim')) %>%
  select(am_sid) %>%
  distinct %>%
  left_join(spp_quad_list, by = 'am_sid')

### This next batch is open Pacific or Indian Ocean species
clipped_spp2 <- am_spp_all_lims %>%
  # filter(long %in% c(77, 185)) %>%
  filter((long == 77  & lim == 'e_lim') | 
         (long == 185 & lim == 'w_lim')) %>%
  select(am_sid) %>%
  distinct %>%
  left_join(spp_quad_list, by = 'am_sid')

write_csv(clipped_spp1, file.path(dir_git, 'clip_fao', 'clipped_se_asia.csv'))
write_csv(clipped_spp2, file.path(dir_git, 'clip_fao', 'clipped_india_pacific.csv'))

@@

``` {r plot_spp, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

clipped_spp1 <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_se_asia.csv'))
clipped_spp2 <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_india_pacific.csv'))

loiczid_raster <- raster('data/rasters/loiczid_raster.tif')

spp_examples <- sample_n(clipped_spp2, 10)

for (i in 1:nrow(spp_examples)) {

  spp <- spp_examples[i, ]
  map_cells <- am_spp_cells %>%
    filter(am_sid == spp$am_sid) %>%
    mutate(am_pres = 1)

@@

  spp_pts <- as.data.frame(rasterToPoints(r_am_spp))

  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
    ggtheme_map + 
    geom_raster(aes(fill = am_pres), alpha = .8, show.legend = FALSE) +
    borders('world', color='gray30', fill='gray40', size = .1) +  # create a layer of borders
    scale_x_continuous(breaks = seq(-180, 180, by = 30)) +
    scale_y_continuous(breaks = seq( -90,  90, by = 30))

  plot(spp_plot)
}

@@

>>>>

library(ggplot2)
library(tmap)
data(World)

library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

@@

am_spp_latlong <- am_spp_latlong %>%
  filter(clongadj >= 60 & clongadj <= 240)

@@

``` {r northmost_band}
### 5N north     - look for spike at 175 W (185E)
am_spp_north <- am_spp_latlong %>%
  filter(centerlat >= 5)

am_spp_north_lims <- am_spp_north %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long <= 70 | long >= 200, NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

@@

am_spp_central_lims <- am_spp_central %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long <= 70 | long >= 200, NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

@@

@@

clipped_spp <- am_spp_central_lims %>%
  # bind_rows(am_spp_north_lims) %>%
  # filter(long %in% c(77, 185)) %>%
  filter(long == 185) %>%
  select(am_sid, lim) %>%
  distinct() %>%
  left_join(spp_quad_list, by = 'am_sid')

write_csv(clipped_spp, file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))

@@

``` {r plot_setup, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

loiczid_raster <- raster('data/rasters/loiczid_raster.tif')
fao_polys <- rgdal::readOGR(dsn = file.path(path.expand(dir_git), 'clip_fao/fao_rgns'),
                            layer = 'fao_rgns2')

assemble_map <- function(map_rast, spp) {
  message('...mapping ', spp)
  map_obj <- tm_shape(map_rast) +
    tm_raster(palette = 'Reds',
              colorNA = NULL,
              title = spp,
              alpha = 1) +
    tm_shape(World) +
      tm_polygons() + 
    tm_layout(basemaps = "Esri.WorldTopoMap", 
              # title.position = 'TOP', 
              legend.outside = TRUE, attr.outside = TRUE)

  return(map_obj)
}
```

``` {r plot_spp_e, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

spp_examples_e <- clipped_spp %>%
  filter(lim == 'w_lim')

rast_list <- vector('list', length = nrow(spp_examples_e))

for (i in 1:nrow(spp_examples_e)) {  # i <- 1

  spp <- spp_examples_e[i, ]
  map_cells <- am_spp_cells %>%
    filter(am_sid == spp$am_sid) %>%
    mutate(am_pres = 1)

@@

  r1 <- r_am_spp %>%
    crop(extent(c(-180, 0, -50, 50)))
  extent(r1) <- c(180, 360, -50, 50)

  r2 <- r_am_spp %>%
    crop(extent(c(0, 180, -50, 50)))

  r3 <- merge(r2, r1)

  new_ext <- extent(c(50, 240, -50, 50))
  r_am_crop <- crop(r3, new_ext)

  rast_list[i] <- r_am_crop
  names(rast_list)[i] <- spp

  spp_plot <- assemble_map(r_am_crop, spp$sciname)
}

mapstack_e <- stack(rast_list) %>%
  calc(sum, na.rm = TRUE)
plot(mapstack_e)
values(mapstack_e)[values(mapstack_e) == 0] <- NA
mapstack_e <- crop(mapstack_e, extent(c(85, 220, -50, 50)))

stack_plot_e <- assemble_map(mapstack_e, 'n_spp') +
  tm_shape(fao_polys) +
  tm_borders()
print(stack_plot_e)

@@

``` {r plot_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

spp_examples_w <- clipped_spp %>%
  filter(lim == 'e_lim') %>%
  sample_n(30)

rast_list <- vector('list', length = nrow(spp_examples_w))

for (i in 1:nrow(spp_examples_w)) {  # i <- 1

  spp <- spp_examples_w[i, ]
  map_cells <- am_spp_cells %>%
    filter(am_sid == spp$am_sid) %>%
    mutate(am_pres = 1)

  r1 <- r_am_spp %>%
    crop(extent(c(-180, 0, -50, 50)))
  extent(r1) <- c(180, 360, -50, 50)

  r2 <- r_am_spp %>%
    crop(extent(c(0, 180, -50, 50)))

  r3 <- merge(r2, r1)

  new_ext <- extent(c(70, 220, -50, 50))
  r_am_crop <- crop(r3, new_ext)

  rast_list[i] <- r_am_crop
  names(rast_list)[i] <- spp

  spp_plot <- assemble_map(r_am_crop, spp$sciname)
}

mapstack_w <- stack(rast_list) %>%
  calc(sum, na.rm = TRUE)
plot(mapstack_w)
values(mapstack_w)[values(mapstack_w) == 0] <- NA
mapstack_w <- crop(mapstack_w, extent(c(85, 220, -50, 50)))

stack_plot_w <- assemble_map(mapstack_w, 'n_spp') +
  tm_shape(fao_polys) +
  tm_borders()

print(stack_plot_w)

``` {r combine_rasts}

stack_plot_both <- 
  tm_shape(mapstack_w) +
    tm_raster(palette = 'Reds',
              colorNA = NULL,
              title = 'n_spp',
              alpha = 1) +
  tm_shape(mapstack_e) +
    tm_raster(palette = 'Blues',
              colorNA = NULL,
              title = 'n_spp') +
  tm_shape(World) +
    tm_polygons() + 
  tm_shape(fao_polys) +
    tm_borders() +
  tm_layout(basemaps = "Esri.WorldTopoMap", 
            # title.position = 'TOP', 
            legend.outside = TRUE, attr.outside = TRUE)

print(stack_plot_both)
```
